---
title: 'ggrrr: Caching functions'
output: html_document
---

# Caching

`ggrrr` supports a super simple passthrough caching layer. This assumes calculations are deterministic, and with a stated set of inputs can 
compute and cache the result.

Setting the caching location should be transparent. The default directory is the user cache directory, but in an analysis project a specific
cache directory may be preferred

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
# Setting a cache in an analysis project might be in a subdirectory of the project
# options("cache.dir"=here::here("cache"))

# If in a package you might be using a package specific cache
# options("cache.dir"=rappdirs::user_cache_dir("my-package"))

# for our purposes a temporary directory is OK.
options("cache.dir"= tempdir())
```

## Caching an item

```{r}
seed = 101
set.seed(seed)
data = tibble(x = abs(rnorm(100,mean = 10, sd=5))) %>% 
  mutate(y = abs(as.integer(x^2*rnorm(100,2,0.5) + x*rnorm(100,5,5) + x*rnorm(100,3,1))))
fit = glm(y ~ x, family = poisson(),data = data)
plot(fit)
ggplot(data, aes(x=x,y=y))+geom_point()
```

