---
title: 'Figure functions'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Figure functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tidyverse)
library(patchwork)
```

```{r setup}
library(ggrrr)
out = outputter(tempdir(check = TRUE),datedFile = FALSE,datedSubdirectory = FALSE)
```

# Opinionated defaults for figures

`ggrrr` sets a default theme and overrides default line sizes and text sizes to base figures around 8pt sans serif labels.

```{r}

ggrrr::gg_pedantic()

plot = ggplot(diamonds, aes(x=carat,y=price,color = color))+geom_point()+annotate("label",x=2,y=10000,label="Hello \u2014 world")+labs(tag = "A")+xlab("Carat\u2082")+ylab("price\u2265")
plot
```

There are utility functions to switch off various bits of the plot

```{r}
plot+gg_hide_X_axis()+gg_hide_Y_axis()+gg_hide_legend()
```

to control the legend size, arrange the plot to fit a narrower space and put a watermark in the plot

```{r}
plot+gg_narrow()+gg_resize_legend(spaceLegend = 1,pointSize = 4,textSize = 12)+gg_watermark("PRELIM")

```

## Tabular data in ggplots

A simple table as a ggplot grob makes it easier to combine plots and tables. This simple table has very little control on the output.

```{r}
tableExample = diamonds %>% 
  mutate(size = cut(carat, breaks=c(-Inf,quantile(carat, probs = c(0.25,0.75)),Inf), labels=c("small","medium","large")) %>% as.factor()) %>%
  group_by(size,cut) %>%
  summarise(
    N = n(),
    depth = sprintf_list("%1.2f (IQR %1.2f \u2014 %1.2f)", quantile(depth,c(0.5,0.25,0.75))),
    price = sprintf_list("%1.0f (IQR %1.0f \u2014 %1.0f)", quantile(price,c(0.5,0.25,0.75)))
  )
  

tbl = tableExample %>% gg_simple_table(font = "Roboto")
p2 = plot+tbl+labs(tag = "B")+patchwork::plot_layout(ncol=1)
p2
```

## Fixed sizes for save

Previewing ggplots in RStudio resizes the plot to fit the windows, scaling the labels independently of the plot. 
This is annoying when you are designing a plot for publication and the label sizes change with the size of the overall image.
`ggrrr` can force ggplot to render it at the size it will be printed at. In doing so it produces both a PNG and PDF output which will be
(very nearly) identical. It also by default saves an .Rdata of the ggplot object allowing later adjustment.

```{r}
f = out("figure1")
p2 %>% gg_save_as(out("figure1"),size = std_size$two_third)
fs::dir_ls(fs::path_dir(f),glob = paste0(f,"*"))
```

In general journals don't like this mixture of figures and tables however simple tables are also good to add as annotations to a plot. 
For this annotating an unwrapped grob is needed e.g.:

```{r}
tableExample2 = diamonds %>% 
  mutate(size = cut(carat, breaks=c(-Inf,quantile(carat, probs = c(0.25,0.75)),Inf), labels=c("small","medium","large")) %>% as.factor()) %>%
  group_by(size) %>%
  summarise(
    N = n(),
    depth = sprintf_list("%1.2f (IQR %1.2f \u2014 %1.2f)", quantile(depth,c(0.5,0.25,0.75))),
    price = sprintf_list("%1.0f (IQR %1.0f \u2014 %1.0f)", quantile(price,c(0.5,0.25,0.75)))
  )

tbl2 = tableExample2 %>% gg_simple_table(font = "Roboto",unwrapped = TRUE)
plot+annotation_custom(tbl2)

```

laying these tables out is tricky, and is worth looking at the `ggpp` package for assistance.

# Scaling tables to fit

A more complex tabular layout can be achieved which auto scales text to fit the specific dimensions provided.

```{r}
# devtools::load_all("~/Git/ggrrr")
# tableExample %>% as.long_format_table(fontName = "Times New Roman",fontSize = 12) %>% glimpse()
tablePlot = tableExample %>% gg_formatted_table(tableWidthInches = 4,font = "Roboto",fontSize = 8) 
tablePlot %>% gg_save_as(tempfile(),maxWidth = 4)
```


