---
title: 'Font functions' 
output: rmarkdown::html_vignette 
vignette: >
  %\VignetteIndexEntry{Font functions} 
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

There are quite a few systems for font management in R. A key part of our work
is trying to make graphs and tables cross platform reproducible, and consistent.
This means dealing with fonts across different platforms.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggrrr)
```

## Transparently ensure a font is available

Instead of specifying a font family as a character in a call to ggplot, we can
use a `check_font()` call which tries to find the font on your system and if not
available downloads it from Google fonts or Brick and registers it with `systemfonts`,
returning the family name. 

ggplot rendering uses `systemfonts` and attempts
to make the output portable for use by embedding webfonts into the svg outputs.
if the `sub` option is `TRUE`, `check_font()` will guess an alternative. 

```{r}
check_font("Helvetica")
```

## Check what fonts are available

Find out what fonts `ggrrr` can find on this system.

```{r}
# list all available fonts
tmp = fonts_available()

# Check if specific fonts are available. Returns
fonts_available(c("Roboto","Arial","Helvetica"))
```

# Download a font from webfont providers

Checks if a local font is available. If not downloads and installs the font and
registers in the `systemfonts` packages. Currently supported are `google` and 
`brick`

```{r}
# clears any registered fonts
reset_fonts()

# if the fonts is named the names are used for the family on this system
check_font(c("Roboto","Arial","Kings","EB Garamond"))

systemfonts::registry_fonts()
```

