---
title: 'Font functions'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Font functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
There are quite a few systems for font management in R. A key part of our work is trying to make graphs and tables cross platform reproducible, and consistent. This means
dealing with fonts across different platforms.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggrrr)
```

## Transparently ensure a font is available

Instead of specifying a font family as a character in a call to ggplot, we can use a `check_font()` call which tries to find the font on your system and if not available downloads it from google fonts and registers it with `sysfonts` and `systemfonts`, returning the family name. 

This also sets ggplot to use showtext as the font provider.

```{r cars}
check_font("Roboto")
```

## Check what fonts are available

Find out what fonts `ggrrr` can find on this system. 

```{r}
# list all available fonts
fonts_available()

# Check if specific fonts are available
fonts_available(c("Roboto","Arial","Helvetica"))
```

# Download a font from google

Downloads and installs the font into both `sysfonts` and `systemfonts` packages. This is generally used internally by `check_font`.

```{r}
google_font("Roboto")
```


