---
title: "Structured lists"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structured lists}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Structured lists are a simple(r) and lighter weight alternative to `vctrs` from
the `tidyverse` package. In `pkgutils` a `struct` is defined as a S3 class based
on a named list of values that is managed as a single entity. E.g. a value and
confidence limits. All `struct`s live in `struct_list`s allowing them to be part
of data frames. This list of S3 classes/lists approach means that `struct`s need
infrastructure to make sure they stay in the same format. 


```{r setup}
# devtools::load_all()
```


The aim of `struct_list` is to be transparent to the user and created in a 
package, but removing the overhead of managing data as lists of lists.

## Constructor

As with `vctrs` the first step is to define a constructor function that performs
any error checking, and handles any vectorisation needed 

```{r}

new_HR = function(mid, lower, upper, p.value) {
  return(struct(
    mid = mid,
    lower = lower,
    upper = upper,
    p.value = p.value,
    .class = "HR"))
}

format.HR = function(x, ...) {
  sprintf("%1.2f [%1.2f \u2014 %1.2f] (P:%s)",x$mid,x$lower,x$upper,scales::pvalue(x$p.value))
}

new_HR(1.2,0.8,1.6,0.8534)

```

Printing as a list of custom S3 object is handled by default using `tibble` and
`pillar` packages.

```{r}

df = tibble::tibble(
  index = 1:5,
  HR = rep(new_HR(1.2,0.8,1.6,0.8534),5)
) 

format(df)
```

and it will also print correctly using `knitr`s default 

```{r}
knitr::kable(df)
# but not directly with huxtable
# huxtable::hux(df %>% dplyr::mutate(HR = format(HR)))
```

As first class lists they can be manipulated with `purrr` functions, for 
extracting or manipulating values. If the 
value we expect back from a purrr::map is a `struct` just like `purrr::map_dbl`
and similar functions we use a `pkgutils::map_struct` function.

```{r}
x = new_HR(1.2,0.8,1.6,0.8534)

x$mid
```


```{r}
df$HR$mid
```
